$title$$if(thanks)$\thanks{$thanks$}$endif$

### 2025-04-19 index.qmd
Add COI statement
Cleaning up

---
title: A Guide to Designing Experiments to Test Statistical Graphics
author:
  - name: Emily Robinson
    dept: Statistics Department
    affiliation: 
    email: erobin17@calpoly.edu
    orcid: 0000-0001-9800-7304
    affiliations: 
      - name: Statistics Department, California Polytechnic State University
        address: Faculty Offices East, Building 25, Room 107D
        city: San Luis Obispo
        state: CA
        postal-code: 93407-0405
  - name: Heike Hofmann
    orcid: 0000-0001-6216-5183
    email: hhofmann4@unl.edu    
    affiliations: 
      - name: Statistics Department, University of Nebraska - Lincoln
        address: 340 Hardin Hall North Wing
        city: Lincoln
        state: NE
        postal-code: 68503
  - name: Susan Vanderplas
    email: svanderplas2@unl.edu
    orcid: 0000-0002-3803-0972
    affiliations: 
      - name: Statistics Department, University of Nebraska - Lincoln
        address: 340 Hardin Hall North Wing
        city: Lincoln
        state: NE
        postal-code: 68503

@@

bibliography: refs.bib
csl: apa.csl

mainfont: Helvetica
monofont: Roboto
number-depth: 2
number-sections: true

@@

header-includes:
   - \usepackage[dvipsnames]{xcolor} % colors
   - \newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
   - \newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
   - \newcommand{\hh}[1]{{\textcolor{Green}{#1}}}

date: last-modified

format:  
  wordcount-pdf:
    keep-tex: true
    fontsize: 10pt
---

@@

```{r lof-density-curves, echo = F, include = T, message = F, warning = F}
#| label: fig-lof-density-curves
#| fig-cap: "Density plot of the lack of fit statistic showing separation of selected difficulty levels: High (obvious curvature), Medium (noticeable curvature), and Low (almost linear). Each density plot is the result of 1000 simulations from a model $y_i = \\alpha\\cdot e^{\\beta\\cdot x_i + \\epsilon_i} + \\theta$, where $\\epsilon \\sim N(0, \\sigma^2)$. $\\alpha$ and $\\theta$ were selected after manipulation of $\\beta$ and $\\sigma$ to ensure that all data generated had similar $y$ ranges so as not to provide visual cues about model differences outside of the plot curvature."
#| out.width: "\\columnwidth"
#| fig.height: 2.5
#| fig.width: 5

@@

```{r, qe2-density-plot, message=FALSE, warning=FALSE, echo = F}
#| label: fig-density-rug
#| fig-cap: "Density of participant estimates for the year (x-value) in which the population reaches 4000 (y-value). Colors are associated to scale - linear (blue) and log (orange) - and vertical lines indicate the true value based on the underlying model equation (black solid) and closest point value based on the simulated data set (black dashed). A jittered rug plot along the $x$-axis shows where participant estimates were made. The plot shows anchoring occurred to the closest point as shown by an increase in density around the dashed line. Density peaks occurred at whole values indicating rounding errors."
#| fig-width: 6
#| fig-height: 3
#| out-width: "\\columnwidth"

@@

>>>>

---
title: A Guide to Designing Experiments to Test Statistical Graphics
authors:
  - id: ear
    name: Emily Robinson
    dept: Statistics Department
    email: erobin17@calpoly.edu
    orcid: 0000-0001-9800-7304
    affiliations: 
    - name: California Polytechnic State University
      department: Statistics Department
      address: Faculty Offices East, Building 25, Room 107D
      city: San Luis Obispo
      state: CA
      postal-code: 93407-0405

  - id: hh
    name: Heike Hofmann
    orcid: 0000-0001-6216-5183
    email: hhofmann4@unl.edu    
    affiliations: 
    - name: University of Nebraska - Lincoln
      department: Statistics Department
      address: 340 Hardin Hall North Wing
      city: Lincoln
      state: NE
      postal-code: 68503

  - id: svp
    name: Susan Vanderplas
    email: svanderplas2@unl.edu
    orcid: 0000-0002-3803-0972
    corresponding: true
    affiliations: 
    - name: University of Nebraska - Lincoln
      department: Statistics Department
      address: 340 Hardin Hall North Wing
      city: Lincoln
      state: NE
      postal-code: 68503

thanks: "The author(s) received no specific funding for this work."

@@

bibliography: refs.bib

@@

header-includes:
   - \usepackage[dvipsnames]{xcolor} % colors
   - \renewcommand{\thefootnote}{\arabic{footnote}}
   - \newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
   - \newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
   - \newcommand{\hh}[1]{{\textcolor{Green}{#1}}}
date: last-modified

format:  
  wordcount-pdf:       
    keep-tex: true
    template-partials:
      - partials/title.tex
      - partials/_print_author.tex
---

@@

```{r, echo = F, include = T, message = F, warning = F}
#| label: fig-lof-density-curves
#| fig-cap: "Density plot of the lack of fit statistic showing separation of selected difficulty levels: High (obvious curvature), Medium (noticeable curvature), and Low (almost linear). Each density plot is the result of 1000 simulations from a model $y_i = \\alpha\\cdot e^{\\beta\\cdot x_i + \\epsilon_i} + \\theta$, where $\\epsilon \\sim N(0, \\sigma^2)$. $\\alpha$ and $\\theta$ were selected after manipulation of $\\beta$ and $\\sigma$ to ensure that all data generated had similar $y$ ranges so as not to provide visual cues about model differences outside of the plot curvature."
#| out.width: "\\columnwidth"
#| fig.height: 2.5
#| fig.width: 5

@@

```{r, message=FALSE, warning=FALSE, echo = F}
#| label: fig-density-rug
#| fig-cap: "Density of participant estimates for the year (x-value) in which the population reaches 4000 (y-value). Colors are associated to scale - linear (blue) and log (orange) - and vertical lines indicate the true value based on the underlying model equation (black solid) and closest point value based on the simulated data set (black dashed). A jittered rug plot along the $x$-axis shows where participant estimates were made. The plot shows anchoring occurred to the closest point as shown by an increase in density around the dashed line. Density peaks occurred at whole values indicating rounding errors."
#| fig-width: 6
#| fig-height: 3
#| out-width: "\\columnwidth"

@@

# Author Statement
The authors have no conflicts of interest or additional funding to declare in this manuscript.

### 2025-04-19 index.tex
Add COI statement
Cleaning up

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
  letterpaper,
  DIV=11,
  numbers=noendperiod]{scrartcl}

@@

\KOMAoption{captions}{tableheading}
\usepackage[dvipsnames]{xcolor} % colors
\newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\hh}[1]{{\textcolor{Green}{#1}}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Guide to Designing Experiments to Test Statistical Graphics},
  pdfauthor={Emily Robinson; Heike Hofmann; Susan Vanderplas},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={Blue},
  pdfcreator={LaTeX via pandoc}}

\title{A Guide to Designing Experiments to Test Statistical Graphics}
\author{Emily Robinson \and Heike Hofmann \and Susan Vanderplas}
\date{2025-03-26}

\begin{document}
\maketitle
\begin{abstract}
In this paper, we discuss considerations and methods for experimentally
testing visualizations. We discuss levels of user engagement with
graphics, common issues when developing a sampling or data generation
model, the importance of pilot testing, and data analysis methods. Along
the way, we also provide recommendations of how to avoid some of the
unique pitfalls of human testing in statistical and visualization
research.
\end{abstract}

@@

>>>>

% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\PassOptionsToPackage{dvipsnames,svgnames,x11names}{xcolor}
%
\documentclass[
  10pt,
]{article}

@@

\usepackage{orcidlink}
\definecolor{mypink}{RGB}{219, 48, 122}
\usepackage[dvipsnames]{xcolor} % colors
\renewcommand{\thefootnote}{\arabic{footnote}}
\newcommand{\ear}[1]{{\textcolor{blue}{#1}}}
\newcommand{\svp}[1]{{\textcolor{RedOrange}{#1}}}
\newcommand{\hh}[1]{{\textcolor{Green}{#1}}}
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\AtBeginDocument{%
\ifdefined\contentsname
  \renewcommand*\contentsname{Table of contents}
\else
  \newcommand\contentsname{Table of contents}
\fi
\ifdefined\listfigurename
  \renewcommand*\listfigurename{List of Figures}
\else
  \newcommand\listfigurename{List of Figures}
\fi
\ifdefined\listtablename
  \renewcommand*\listtablename{List of Tables}
\else
  \newcommand\listtablename{List of Tables}
\fi
\ifdefined\figurename
  \renewcommand*\figurename{Figure}
\else
  \newcommand\figurename{Figure}
\fi
\ifdefined\tablename
  \renewcommand*\tablename{Table}
\else
  \newcommand\tablename{Table}
\fi
}
\@ifpackageloaded{float}{}{\usepackage{float}}
\floatstyle{ruled}
\@ifundefined{c@chapter}{\newfloat{codelisting}{h}{lop}}{\newfloat{codelisting}{h}{lop}[chapter]}
\floatname{codelisting}{Listing}
\newcommand*\listoflistings{\listof{codelisting}{List of Listings}}
\makeatother
\makeatletter
\makeatother
\makeatletter
\@ifpackageloaded{caption}{}{\usepackage{caption}}
\@ifpackageloaded{subcaption}{}{\usepackage{subcaption}}
\makeatother

@@

\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same} % disable monospaced font for URLs
\hypersetup{
  pdftitle={A Guide to Designing Experiments to Test Statistical Graphics},
  colorlinks=true,
  linkcolor={blue},
  filecolor={Maroon},
  citecolor={Blue},
  urlcolor={red},
  pdfcreator={LaTeX via pandoc}}

\title{A Guide to Designing Experiments to Test Statistical
Graphics\thanks{The author(s) received no specific funding for this
work.}}

%% Author information
\author{Emily
Robinson\footnote{Email: erobin17@calpoly.edu}~\orcidlink{0000-0001-9800-7304}\\{\footnotesize Statistics
Department}, {\footnotesize California Polytechnic State
University}\\Heike
Hofmann\footnote{Email: hhofmann4@unl.edu}~\orcidlink{0000-0001-6216-5183}\\{\footnotesize Statistics
Department}, {\footnotesize University of Nebraska - Lincoln}\\Susan
Vanderplas\footnote{Email: svanderplas2@unl.edu}~\orcidlink{0000-0002-3803-0972}\\{\footnotesize Statistics
Department}, {\footnotesize University of Nebraska - Lincoln}\\}
\begin{document}
\maketitle
\begin{abstract}
In this paper, we discuss considerations and methods for experimentally
testing visualizations. We discuss levels of user engagement with
graphics, common issues when developing a sampling or data generation
model, the importance of pilot testing, and data analysis methods. Along
the way, we also provide recommendations of how to avoid some of the
unique pitfalls of human testing in statistical and visualization
research.
\end{abstract}

@@

\section{Author Statement}\label{author-statement}

The authors have no conflicts of interest or additional funding to
declare in this manuscript.

### 2025-04-19 partials/title.tex
Cleaning up

$if(title)$
  \title{$title$}
$endif$
$if(subtitle)$
  \subtitle{$subtitle$}
$endif$

\author{%
$for(author-list)$%
  $it.name$\thanks{email: $it.email$}\\\scriptsize $it.dept$\\\scriptsize $it.affiliation$%
  $sep$\and%
$endfor$%
}

$if(date)$
  \date{$date$}
$endif$

$if(teaserimg)$
\teaser{
  \centering
  \includegraphics[width=\linewidth]{$teaser-img$}
  \caption{$teaser-caption$}
  \label{fig:teaser}
}
$endif$

$if(abstract)$
\abstract{$abstract$}
$endif$

\vgtccategory{$category$}

$if(ieee-keywords)$
\CCScatlist{
$for(ieee-keywords)$
$it$
$sep$;
$endfor$
}
$endif$

>>>>

$-- Provides configuration of document metadata for writing the title block.
$-- Note that in addition to these templates and partials, Quarto will also make normalized authors and affiliations data available to the template,
$-- making is easy to write custom title blocks against a standard schema.
$--
$-- %%%% TODO %%%%%
$-- Customize is needed, like below for printing the authors. Otherwise remove this partials to use Quarto default one.
$-- %%%%%%%%%%%%%%%%

$if(title)$
\title{$title$$if(thanks)$\thanks{$thanks$}$endif$}
$endif$
$if(subtitle)$
\usepackage{etoolbox}
\makeatletter
\providecommand{\subtitle}[1]{% add subtitle to \maketitle
  \apptocmd{\@title}{\par {\large #1 \par}}{}{}
}
\makeatother
\subtitle{$subtitle$}
$endif$
$-- This supports Quarto's authors normalization

%% Author information
\author{$for(by-author)$$_print_author.tex()$\\$endfor$}
